The evolution of gene expression in the term placenta of viviparous mammals

### 2015-10-21 manuscript/manuscript.Rnw
clean up bonobo results section some more

The top 50 expressed genes as measured by \ac{FPKM} in the term placenta of
the bonobo are shown in \fref{tab:top50_bonobo_placenta} with a list of all genes in
Supplemental Table X. Expression in the bonobo placenta differs
somewhat from that of the human, as evidenced by the makeup of the top
50 genes (Table 3). The highest expression (\ac{FPKM}=%
\Sexpr{gene.expression("pan paniscus","PRG2")}%
%
) in bonobo is \ac{PRG2}, which is involved in insulin
growth factor signaling and angiogenesis during pregnancy \cite{Weyer.ea2011:Placentalregulationofpeptide}. \ac{PRG2} is also expressed at relatively high levels in the
human placenta (\ac{FPKM}=\Sexpr{gene.expression("homo sapiens","PRG2")}; Supplemental Table X), although at
comparatively much lower levels than the bonobo. A bonobo gene closely
related to the chorionic somatomammotropins (Q95MK7\_PANTR) is also
expressed at high levels (\ac{FPKM}=1232.56; \fref{tab:top50_bonobo_placenta}). Two other Pan genes
included within this family (ENSPTRG00000030805 which shows similarity
to GH2 and ENSPTRG00000034560) show expression levels of \ac{FPKM}>89. Six
pregnancy-specific beta-1-glycoprotein genes were expressed in the
bonobo placenta with \ac{FPKM} ranging from 20.885-167.961 (Supplemental
Table X). Placental expression of the genes involved in estrogen and
progesterone production (CYP19A1, HSD3B1 respectively) show lower
expression in the bonobo as compared to humans (CYP19A1 \ac{FPKM}= 48.6877;
HSD3B1 \ac{FPKM}= 54.5301; Table 8). In addition, CGA is expressed in the
bonobo placenta (\ac{FPKM}=202.629), but at much lower levels than humans
as well (Table 8). Two Ensembl-annotated beta subunits of the
chorionic gonadotropin genes are expressed in bonobo, one related to
CGB and another related to CGB7; both are not highly expressed (Table
8). Among the other top 50 expressed genes, there is high expression
of ribosomal proteins (e.g., RPL18, RPL23), keratins (e.g., KRT7,
KRT19), heat-shock proteins (e.g., HSPA5, HSPA1A), serpins (SERPINE1,
SERPINE2), insulin growth factor 2 (IGF2) and the hormone
adrenomedullin (ADM).

>>>>

The top 50 expressed genes as measured by \ac{FPKM} in the term
placenta of the bonobo are shown in \fref{tab:top50_bonobo_placenta}
with a list of all genes in Supplemental
\fref{tab:all_pan_paniscus_placenta}. Expression in the bonobo
placenta differs somewhat from that of the human, as evidenced by the
makeup of the top 50 genes (\fref{tab:top50_bonobo_placenta}). The highest expression
(\ac{FPKM}=%
\Sexpr{gene.expression("pan paniscus","PRG2")}%
%
) in bonobo is \textit{PRG2}, which is involved in insulin
growth factor signaling and angiogenesis during pregnancy \cite{Weyer.ea2011:Placentalregulationofpeptide}. \textit{PRG2} is also expressed at relatively high levels in the
human placenta (\ac{FPKM}=\Sexpr{gene.expression("homo sapiens","PRG2")}; Supplemental \fref{tab:all_homo_sapiens_placenta}), although at
comparatively much lower levels than the bonobo. A bonobo gene closely
related to the chorionic somatomammotropins (Q95MK7\_PANTR) is also
expressed at high levels (\ac{FPKM}=1232.56; \fref{tab:top50_bonobo_placenta}). Two other Pan genes
included within this family (ENSPTRG00000030805 which shows similarity
to GH2 and ENSPTRG00000034560) show expression levels of \ac{FPKM}>89. Six
pregnancy-specific beta-1-glycoprotein genes were expressed in the
bonobo placenta with \ac{FPKM} ranging from 20.885-167.961 (Supplemental
Table X). Placental expression of the genes involved in estrogen and
progesterone production (CYP19A1, HSD3B1 respectively) show lower
expression in the bonobo as compared to humans (CYP19A1 \ac{FPKM}= 48.6877;
HSD3B1 \ac{FPKM}= 54.5301; Table 8). In addition, CGA is expressed in the
bonobo placenta (\ac{FPKM}=202.629), but at much lower levels than humans
as well (Table 8). Two Ensembl-annotated beta subunits of the
chorionic gonadotropin genes are expressed in bonobo, one related to
CGB and another related to CGB7; both are not highly expressed (Table
8). Among the other top 50 expressed genes, there is high expression
of ribosomal proteins (e.g., RPL18, RPL23), keratins (e.g., KRT7,
KRT19), heat-shock proteins (e.g., HSPA5, HSPA1A), serpins (SERPINE1,
SERPINE2), insulin growth factor 2 (IGF2) and the hormone
adrenomedullin (ADM).
